---
title: "CB 7 Collision Report"
author: "Joyce Robbins"
date: "October 28, 2015"
output: word_document
---
```{r echo = FALSE, message = FALSE}
# preliminary steps
library(dplyr)
library(lubridate)
library(RColorBrewer)

dat <- read.csv("CB7-10023-10024-10025-10069.csv", stringsAsFactors = FALSE, header = TRUE)
# d$IK is greater than 0 if someone is injured or killed 
dat$IK <- dat[,11] + dat[,12] 
dat$DATE <- as.Date(dat$DATE, "%m/ %d/ %Y")
yearlab = c("2013", "2014", "2015 (to date)")

newdata <- dat %>%
 filter (IK > 0) %>%
 mutate(year = year(DATE), monthnum = month(DATE), month = month(DATE, label=T)) %>%
 filter (year >= 2013)
```



## Data Source
https://data.cityofnewyork.us/NYC-BigApps/CB7-10023-10024-10025-10069-/3deq-ujhm

## Number of Persons Injured

```{r echo = FALSE}
injured <- newdata %>% group_by(year)  %>% 
   select(contains ("INJURED")) %>% 
   summarise_each(funs(sum)) 
rownames(injured) <- injured$year
injured <- select(injured, -year)
labels <- c("Total", "Pedestrians", "Cyclists", "Motorists")
barplot(as.matrix(injured), ylim = c(0, max(injured)+100), beside = TRUE, names = labels, legend.text = yearlab)
```
The number of injuries went down from 2013 to 2014...

## Number of Persons Killed

```{r echo = FALSE}
killed <- newdata %>% group_by(year)  %>% select(contains ("KILLED")) %>% summarise_each(funs(sum))
rownames(killed) <- killed$year
killed <- select(killed, -year)
labels <- c("Total", "Pedestrians", "Cyclists", "Motorists")
barplot(as.matrix(killed), ylim = c(0, max(killed)+10), beside = TRUE, names = labels, legend.text = yearlab)
```

... but the number killed went up from `r killed[which(rownames(killed)=="2013"),1]` to `r killed[which(rownames(killed)=="2014"),1]` from 2013 to 2014.  

## Monthly Pattern of Injuries/Deaths
``` {r echo = FALSE}
monthdata <- newdata %>% group_by(year, month) %>%
   arrange(year, monthnum) %>%
   summarise(IK = sum(IK))
ts <- ts(monthdata$IK, start = c(2013,1), frequency = 12)

monthplot(ts, ylab = "# of injuries/deaths")

md <- ungroup(monthdata)
h <- head(arrange(md, desc(IK)), 5)
t <- tail(arrange(md, desc(IK)), 5)

```

## Observations
June is consistently high:  
```{r echo=FALSE}
june <- which(monthdata$month == "Jun")
cat(paste("June",monthdata$year[june], "--",monthdata$IK[june], "injured/killed", collapse = "\n"))
```

Otherwise, no clear pattern emerges. There are no repeated months in either the top five or bottom five for injured/killed, and `r length(intersect(h$month,t$month))` months appear in both lists: `r intersect(h$month,t$month)`.  

Top five: 
``` {r echo = FALSE}
cat(paste0(h$month, " ", h$year, " (" ,h$IK, ")", collapse="\n"))
```
Bottom five:  
``` {r echo = FALSE}
cat(paste0(t$month, " ", t$year, " (" ,t$IK, ")", collapse="\n"))
``` 
 
## Location
Plotting (longitude, latitude) shows that injured/killed are distributed throughout the community board district:

```{r echo=FALSE}
library(OpenStreetMap)
library(rgdal)
library(rJava)
map <- openmap(c(40.8277,-74.0257), c(40.7552,-73.8513), type = "osm")
#map <- openmap(c(43.611522,1.428533), c(43.598903,1.461059), type = "osm")
plot(map)


 #openmap(upperLeft, lowerRight, zoom = NULL, type = c("osm", "osm-bw",
 #     "maptoolkit-topo", "waze", "mapquest", "mapquest-aerial", "bing",
 #     "stamen-toner", "stamen-terrain", "stamen-watercolor", "osm-german",
 #     "osm-wanderreitkarte", "mapbox", "esri", "esri-topo", "nps", "apple-iphoto",
 #     "skobbler", "cloudmade-<id>", "hillshade", "opencyclemap", "osm-transport",
 #     "osm-public-transport", "osm-bbike", "osm-bbike-german"), minNumTiles = 9L,
 #     mergeTiles = TRUE)

t <- tally(group_by(newdata, LONGITUDE, LATITUDE))
c <- brewer.pal (n = 9, name = "Blues")
c <- c[5:9]
d <- rbind(c,c,c,c)
dim(d) <- c(20,1)
d <- d[,1]
colors <- d
points(t$LONGITUDE,t$LATITUDE, asp = 1, cex = .2*t$n, col = colors[t$n])

most <- which (newdata$LONGITUDE == t$LONGITUDE[which.max(t$n)])
```

## Intersection with the highest # of injured/killed  

Intersection: `r newdata$ON.STREET.NAME[most[1]]` and `r newdata$CROSS.STREET.NAME[most[1]]`  
Collisions by month: 
```{r echo=FALSE}
tally(group_by(newdata[most,], year, month))
```


